---
title: "Bayesian_NIMBLE_Assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nimble)
library(rjags)
library(lubridate)
library(tidybayes)
library(tidyverse)
```
Very much a work in progress
Made by ASL 16 Mar 2021

First attempt at the nimble model for all sites. Currently sites are different groups in the hierarchical model. This takes a LONG time to run on my computer

MEL: Plots or sites? I had no luck getting a site effect to converge so am only doing plots. Which I think is what you were doing?? Ugh plots, sites, plots, sites, plites, sots...
```{r}
ticks <- read_csv("../0_Data_files/ticks-targets.csv.gz")
drivers <- read.csv("../0_Data_files/PRISM_data/combined_prism_drivers.csv")

#NOTE: this does not include all data right now! Only days with driver data available
data = drivers%>%
  rename(time = Date,
         siteID = Site)%>%
  mutate(time = as.Date(time))%>%
  left_join(ticks)

#KEY CHANGE: limiting to only timepoints when we have observations - since we are not currently using state-space anyway, no need to include NA observations at this time, and it facilitates convergence
data1 <- data %>%
  filter(year(time) %in% c(2016:2018)) %>%
  mutate(NewEpiWeek = epiweek(time)) %>%
  filter(NewEpiWeek %in% c(09:49)) %>%
  mutate(NewEpiWeek = sprintf("%02d", NewEpiWeek),
         Year = year(time)) %>%
  gather(ambloyomma_americanum:ixodes_scapularis, key = "Species", value = "Count") %>%
  select(-Plot) %>%
  rename(Site = siteID,
         Date = time,
         Plot = plotID,
         Temp_C = tmean..degrees.C.,
         Precip_mm = ppt..mm.) %>%
  select(Site, Year, Date, Plot, Count, Species, Temp_C, Precip_mm) %>%
  filter(!is.na(Count)) %>%
  arrange(Date)

# Set up indices
spp <- as.numeric(factor(data1$Species))
yrs <- as.numeric(factor(data1$Year))
sites <- as.numeric(factor(data1$Site))
plots <- as.numeric(factor(data1$Plot))

# Plot counts vs. temp and precip
ggplot(data = data1, aes(x = Temp_C, y = log(Count), group = Species, color = Species))+
  geom_point(size = 2)+
  theme_classic()+
  geom_smooth()

ggplot(data = data1, aes(x = Precip_mm, y = log(Count), group = Species, color = Species))+
  geom_point(size = 2)+
  theme_classic()

```

```{r}
tick_model <- nimbleCode({

  # Priors
  
  #random effects (or fixed effects? I am not sure how you would code that up differently)
  mu_plot ~ dnorm(0, 100)
  mu_sp ~ dnorm(0, 100)
  sd_plot ~ dgamma(0.01, 0.01)
  sd_sp ~ dgamma(0.01, 0.01)
  
  for(l in 1:n_plots){
    plot[l] ~ dnorm(mu_plot, sd = sd_plot)
  }
  
  for(m in 1:n_spp){ 
    sp[m] ~ dnorm(mu_sp, sd = sd_sp)
  }

  #Loop through data points
  for(i in 1:n){
    
      # Process model
      log(pred[i]) <- plot[plots[i]]*x[i] + sp[spp[i]]  
      
      # Data model
      y[i]  ~ dpois(pred[i])
  }

})

constants <- list(n = length(data1$Date),
                  n_spp = length(unique(data1$Species)),
                  n_plots = length(unique(data1$Plot)),
                  spp = spp,
                  plots = plots)

data <- list(x = data1$Temp_C,
             y = data1$Count )

nchain <- 3
inits <- list()
for(i in 1:nchain){
  inits[[i]] <- list(mu_plot = rnorm(1, 1, 1),
                     mu_sp = rnorm(1, 1, 1),
                     sd_sp = rgamma(1, 1, 1),
                     sd_plot = rgamma(1, 1, 1))
}

model$initializeInfo()

nimble.out <- nimbleMCMC(code = tick_model,
                           data = data,
                           inits = inits,
                           constants = constants,
                           monitors = c("sd_sp","sd_plot","mu_plot","mu_sp"),
                           niter = 10000,
                           nchains = 3,
                           samplesAsCodaMCMC = TRUE)

plot(nimble.out) #BOOM
gelman.diag(nimble.out)  ## BOOM!! :-)

#Haven't gotten to this point yet
burnin <- 2000                               
nimble.burn <- window(nimble.out, start=burnin)

plot(nimble.burn) #
effectiveSize(nimble.burn)
gelman.diag(nimble.burn)  ## determine convergence


##STOPPED ADAPTING HERE
chain <- nimble.burn %>%
  tidybayes::spread_draws(theta1, theta2, theta4, sd_data)

pred_function <- function(p1, temp, group_i){
  p1[group_i] * temp[i]
}
```



Added by Sarah 22 Mar 21:
```{r}
###Goals this week : find reliable priors and dynamic model to mess with in the nimble framework. You gotta represent a couple of different sites with different populations. Historically some of the sites have had zero count and some have had a far number. Try to figure out 


unique(ticks$siteID)


ticks%>% filter(siteID == "TALL") %>% ggplot(aes(x= epiWeek, y= ixodes_scapularis, color= plotID))+
  geom_point()

ticks%>% filter(siteID == "TALL") %>% ggplot(aes(x= epiWeek, y= ambloyomma_americanum, color= plotID))+
  geom_point()

ticks_drivers<- full_join(ticks, drivers, by= "plotID")


ticks_drivers %>% filter(siteID == "TALL") %>% ggplot(aes(x=  ambloyomma_americanum, y= tmean..degrees.C., color= plotID))+
  geom_point()

```

