---
title: "Bayesian Tick Forecast Mar 30 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nimble)
library(rjags)
library(lubridate)
library(tidybayes)
library(tidyverse)
```

## Section 1: Load and format data
Data we are using include tick data targets from the EFI forecasting challenge, as well as climate data from PRISM (https://prism.oregonstate.edu/) as our only driver. In the section below we load in data, combine the two dataframes, omit NAs, and set up indeces for random effects
```{r}
ticks <- read_csv("../0_Data_files/ticks-targets.csv.gz")
drivers <- read.csv("../0_Data_files/PRISM_data/combined_prism_drivers.csv")

#NOTE: this does not include all tick data right now! Only days with driver data available
data = drivers%>%
  rename(time = Date,
         siteID = Site)%>%
  mutate(time = as.Date(time))%>%
  left_join(ticks)

#KEY CHANGE: limiting to only timepoints when we have observations - since we are not currently using state-space anyway, no need to include NA observations at this time, and it facilitates convergence
full_data <- data %>%
  filter(year(time) %in% c(2016:2018)) %>%
  mutate(NewEpiWeek = epiweek(time)) %>%
  filter(NewEpiWeek %in% c(09:49)) %>%
  mutate(NewEpiWeek = sprintf("%02d", NewEpiWeek),
         Year = year(time)) %>%
  gather(ambloyomma_americanum:ixodes_scapularis, key = "Species", value = "Count") %>%
  select(-Plot) %>%
  rename(Site = siteID,
         Date = time,
         Plot = plotID,
         Temp_C = tmean..degrees.C.,
         Precip_mm = ppt..mm.) %>%
  select(Site, Year, Date, Plot, Count, Species, Temp_C, Precip_mm) %>%
  filter(!is.na(Count)) %>%
  arrange(Date)
```

## Section 2: Set up model
Right now we are using a simple model that relates tick measurements to temperature, with a random effect for plot
```{r}
tick_model <- nimbleCode({

  # Priors
  #process error
  sd_proc ~ dgamma(0.01, 0.01)
  
  #random effects 
  mu_plot ~ dnorm(0, 100)
  #mu_site ~ dnorm(0, 100)
  sd_plot ~ dgamma(0.01, 0.01)
  #sd_site ~ dgamma(0.01, 0.01)
  
  for(l in 1:n_plots){
    plot[l] ~ dnorm(mu_plot, sd = sd_plot)
  }
  
  #for(l in 1:n_sites){
  #  site[l] ~ dnorm(mu_site, sd = sd_site)
  #}

  #Loop through data points
  for(i in 1:n){
    
      # Process model
      log(mu[i]) <- plot[plots[i]]*x[i] #+ site[sites[i]]*x[i]
      log(pred[i]) ~ dnorm(log(mu[i]), sd = sd_proc)
      
      # Data model
      y[i]  ~ dpois(pred[i])
  }

})
```

## Section 3: Run model and plot results
```{r}
species_to_run = c("ambloyomma_americanum","ixodes_scapularis")

for(species in species_to_run){
  
  data1<- full_data%>%
    filter(Species == species)
  
  # Set up indices
  yrs <- as.numeric(factor(data1$Year))
  sites <- as.numeric(factor(data1$Site))
  plots <- as.numeric(factor(data1$Plot))
  
  constants <- list(n = length(data1$Date),
                    n_spp = length(unique(data1$Species)),
                    n_plots = length(unique(data1$Plot)),
                    #n_sites = length(unique(data1$Site)),
                    #sites = sites,
                    plots = plots)
  
  data <- list(x = data1$Temp_C,
               y = data1$Count )
  
  nchain <- 3
  inits <- list()
  for(i in 1:nchain){
    inits[[i]] <- list(mu_plot = rnorm(1, 1, 1),
                       #mu_site = rnorm(1, 1, 1),
                       sd_plot = rgamma(1, 1, 1),
                       #sd_site = rgamma(1, 1, 1),
                       sd_proc = rgamma(1, 1, 1))
  }
  
  nimble.out <- nimbleMCMC(code = tick_model,
                             data = data,
                             inits = inits,
                             constants = constants,
                             monitors = c("sd_plot","mu_plot","sd_proc", "pred"), #"sd_site", "mu_site"
                             niter = 10000,
                             nchains = 3,
                             samplesAsCodaMCMC = TRUE)
  
  plot(nimble.out[, c("sd_plot","mu_plot","sd_proc")]) #BOOM
  gelman.diag(nimble.out[, c("sd_plot","mu_plot","sd_proc")])  ## BOOM!! :-)
  
  burnin <- 1000                               
  nimble.burn <- window(nimble.out, start=burnin)
  
  plot(nimble.burn[, c("sd_plot","mu_plot","sd_proc")]) #
  effectiveSize(nimble.burn[, c("sd_plot","mu_plot","sd_proc")])
  gelman.diag(nimble.burn[, c("sd_plot","mu_plot","sd_proc")])  ## determine convergence
  
  data1$row <- as.numeric(row.names(data1))
  
  chain <- nimble.burn %>%
    tidybayes::spread_draws(pred[row])%>%
    group_by(row) %>% 
    summarise(mean = mean(pred),
              upper = quantile(pred, 0.975),
              lower = quantile(pred, 0.025),.groups = "drop")
  
  to_plot = chain%>%
    left_join(data1, by = "row")%>%
    mutate(doy = yday(Date))
  
  sites <- unique(to_plot$Site)
  for (i in 1:length(sites)){
    current_site <- subset(to_plot, Site == sites[i])
    if(nrow(current_site) > 0){
      current_site = current_site%>%
        ggplot(aes(x = doy, color = Plot, fill = Plot))+
        geom_point(aes(y = Count), size = 2)+
        geom_line(aes(y = mean))+
        geom_ribbon(aes(ymin = lower, ymax = upper, color = Plot, fill = Plot), alpha = 0.2) +
        ggtitle(paste0(species))+
        facet_grid(vars(Year), scales = "free_x")+
        theme_bw()
      print(current_site)
    }
  }
}


```
