---
title: "Bayesian Tick Forecast Mar 30 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nimble)
library(rjags)
library(lubridate)
library(tidybayes)
library(tidyverse)
```

## Section 1: Load and format data
Data we are using include tick data targets from the EFI forecasting challenge, as well as climate data from PRISM (https://prism.oregonstate.edu/) as our only driver. In the section below we load in data, combine the two dataframes, omit NAs, and set up indeces for random effects
```{r}
ticks <- read_csv("../0_Data_files/ticks-targets.csv.gz")
drivers <- read.csv("../0_Data_files/PRISM_data/combined_prism_drivers.csv")

#NOTE: this does not include all tick data right now! Only days with driver data available
data = drivers%>%
  rename(time = Date,
         siteID = Site)%>%
  mutate(time = as.Date(time))%>%
  left_join(ticks)

#KEY CHANGE: limiting to only timepoints when we have observations - since we are not currently using state-space anyway, no need to include NA observations at this time, and it facilitates convergence
data1 <- data %>%
  filter(year(time) %in% c(2016:2018)) %>%
  mutate(NewEpiWeek = epiweek(time)) %>%
  filter(NewEpiWeek %in% c(09:49)) %>%
  mutate(NewEpiWeek = sprintf("%02d", NewEpiWeek),
         Year = year(time)) %>%
  gather(ambloyomma_americanum:ixodes_scapularis, key = "Species", value = "Count") %>%
  select(-Plot) %>%
  rename(Site = siteID,
         Date = time,
         Plot = plotID,
         Temp_C = tmean..degrees.C.,
         Precip_mm = ppt..mm.) %>%
  select(Site, Year, Date, Plot, Count, Species, Temp_C, Precip_mm) %>%
  filter(!is.na(Count)) %>%
  arrange(Date)

# Set up indices
spp <- as.numeric(factor(data1$Species))
yrs <- as.numeric(factor(data1$Year))
sites <- as.numeric(factor(data1$Site))
plots <- as.numeric(factor(data1$Plot))
```


## Section 2: Explore driver data
Right now we are using temperature as our only driver, but it looks as if precipitation may also be a useful predictor to incorporate in the future
```{r}
# Plot counts vs. temp and precip
ggplot(data = data1, aes(x = Temp_C, y = log(Count), group = Species, color = Species))+
  geom_point(size = 2)+
  theme_classic()+
  geom_smooth()

ggplot(data = data1, aes(x = Precip_mm, y = log(Count), group = Species, color = Species))+
  geom_point(size = 2)+
  theme_classic()
```


## Section 3: Run model and check convergence
Right now we are using a simple model that relates tick measurements to temperature, with random effects for plot and species
```{r}
tick_model <- nimbleCode({

  # Priors
  #process error
  sd_proc ~ dgamma(0.01, 0.01)
  
  #random effects 
  mu_plot ~ dnorm(0, 100)
  mu_sp ~ dnorm(0, 100)
  #mu_site ~ dnorm(0, 100)
  sd_plot ~ dgamma(0.01, 0.01)
  sd_sp ~ dgamma(0.01, 0.01)
  #sd_site ~ dgamma(0.01, 0.01)
  
  for(l in 1:n_plots){
    plot[l] ~ dnorm(mu_plot, sd = sd_plot)
  }
  
  #for(l in 1:n_sites){
  #  site[l] ~ dnorm(mu_site, sd = sd_site)
  #}
  
  for(m in 1:n_spp){ 
    sp[m] ~ dnorm(mu_sp, sd = sd_sp)
  }

  #Loop through data points
  for(i in 1:n){
    
      # Process model
      log(mu[i]) <- plot[plots[i]]*x[i]  + sp[spp[i]]  #+ site[sites[i]]*x[i]
      log(pred[i]) ~ dnorm(log(mu[i]), sd = sd_proc)
      
      # Data model
      y[i]  ~ dpois(pred[i])
  }

})

constants <- list(n = length(data1$Date),
                  n_spp = length(unique(data1$Species)),
                  n_plots = length(unique(data1$Plot)),
                  #n_sites = length(unique(data1$Site)),
                  spp = spp,
                  #sites = sites,
                  plots = plots)

data <- list(x = data1$Temp_C,
             y = data1$Count )

nchain <- 3
inits <- list()
for(i in 1:nchain){
  inits[[i]] <- list(mu_plot = rnorm(1, 1, 1),
                     mu_sp = rnorm(1, 1, 1),
                     #mu_site = rnorm(1, 1, 1),
                     sd_sp = rgamma(1, 1, 1),
                     sd_plot = rgamma(1, 1, 1),
                     #sd_site = rgamma(1, 1, 1),
                     sd_proc = rgamma(1, 1, 1))
}

nimble.out <- nimbleMCMC(code = tick_model,
                           data = data,
                           inits = inits,
                           constants = constants,
                           monitors = c("sd_sp","sd_plot","mu_plot","mu_sp","sd_proc", "pred"), #"sd_site", "mu_site"
                           niter = 10000,
                           nchains = 3,
                           samplesAsCodaMCMC = TRUE)

plot(nimble.out[, c("sd_sp","sd_plot","mu_plot","mu_sp","sd_proc")]) #BOOM
gelman.diag(nimble.out[, c("sd_sp","sd_plot","mu_plot","mu_sp","sd_proc")])  ## BOOM!! :-)

burnin <- 1000                               
nimble.burn <- window(nimble.out, start=burnin)

plot(nimble.burn[, c("sd_sp","sd_plot","mu_plot","mu_sp","sd_proc")]) #
effectiveSize(nimble.burn[, c("sd_sp","sd_plot","mu_plot","mu_sp","sd_proc")])
gelman.diag(nimble.burn[, c("sd_sp","sd_plot","mu_plot","mu_sp","sd_proc")])  ## determine convergence
```

## Section 4: Plot results
This section takes the model output and plots modeled vs observed tick density across plots and sites for both species. The model does well at recreating observed patterns within the data used to fit the model, but has not been tested on other data
```{r}
data1$row <- as.numeric(row.names(data1))

chain <- nimble.burn %>%
  tidybayes::spread_draws(pred[row])%>%
  group_by(row) %>% 
  summarise(mean = mean(pred),
            upper = quantile(pred, 0.975),
            lower = quantile(pred, 0.025),.groups = "drop")

to_plot = chain%>%
  left_join(data1, by = "row")%>%
  mutate(doy = yday(Date))

sites <- unique(to_plot$Site)
for (i in 1:length(sites)){
  current_site <- subset(to_plot, Site == sites[i])
  
  aa <- current_site%>%
    filter(Species == "ambloyomma_americanum")
  if(nrow(aa) > 0){
    aa = aa%>%
      ggplot(aes(x = doy, color = Plot, fill = Plot))+
      geom_point(aes(y = Count), size = 2)+
      geom_line(aes(y = mean))+
      geom_ribbon(aes(ymin = lower, ymax = upper, color = Plot, fill = Plot), alpha = 0.2) +
      ggtitle("A. americanum")+
      facet_grid(vars(Year), scales = "free_x")+
      theme_bw()
    print(aa)
  }
  
  is <- current_site%>%
    filter(Species == "ixodes_scapularis")
  if(nrow(is)>0){
    is = is%>%
      ggplot(aes(x = doy, y = Count, color = Plot, fill = Plot))+
      geom_point(aes(y = Count), size = 2)+
      geom_line(aes(y = mean))+
      geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
      ggtitle("I. scapularis")+
      facet_grid(vars(Year), scales = "free_x")+
      theme_bw()
    print(is)
  }
}
```
